\frb: Multi-Telescope Burst Properties and Implications for FRB Population

### 2016-11-28 repeater-vladata.tex
worked in new lum disn figure. luminosities in table. added more red text to get input

This paper is part of a series that presents the first localization and host identification of an FRB (CITE: LOC, OPT, EVN, RADIO CPART, XRAY, EFF, WEIRD). \frb, also known as the ``repeating FRB'', was first detected in November 2012 by the Arecibo Observatory \citep{2014ApJ...790..101S}. In mid 2015, new Arecibo observations revealed a series of bursts at the same DM and sky position demonstrating that FRBs are capable of repetition \citep{2016Natur.531..202S}. Beginning in August of 2015, we made the first of nine detections of \frb\ with the Very Large Array (CITE: LOC) and localized it with a precision of 0.1\arcsec. Deep radio and optical observing shows that \frb\ is unambiguously associated with a persistent radio and optical source at a redshift of 0.193 (CITE: OPT, EVN).

\frb\ has now been localized three orders of magnitude better than any other FRB and placed at a cosmological distance. Its lookback and luminosity distances are 746 and 972 Mpc \citep{planck15}, which are orders of magnitude larger than any other millisecond transient. This shows that FRBs:
\begin{itemize}
 \item are extremely luminous, 
 \item have a significant DM contribution from the IGM, and
 \item can be used to probe the IGM and their host galaxy.
\end{itemize}
The promise implied by the first reported FRB \citep{2007Sci...318..777L} is now being realized.

The cosmological distance for \frb\ could have wide-ranging implications for the FRB population as a whole. However, it has not been demonstrated that \frb\ is representative of the overall FRB population. In fact, the repetition of its bursts is unique among all FRBs \citep{2015MNRAS.454..457P}, so it is natural to ask whether \frb\ is representative. An important first step is to demonstrate that the properties of \frb\ are consistent with the significant body of facts for the overall population \citep{2015MNRAS.451.3278M, 2016MPLA...3130013K}. The repeating nature of \frb\ provides us with several statistical tests we can use to test this connection.

@@

Given that FRBs are now known to be useful probes of the IGM, there is strong justification for continuing searches for new detections and localizations. The relatively faint counterpart to \frb\ argues that direct localization of the radio burst will continue to be the best way to find optical hosts to measure distances. Our multi-telescope constraints on burst spectra, measurement of host properties, burst rate estimates, and other properties will inform new strategies for finding FRBs...

@@

\begin{figure*}[t]
\begin{center}
\includegraphics[width=2\columnwidth]{timeline}
\caption{Summary of observing coverage and detections of \frb\ during the multi-telescope observing campaign. Symbols show days with observations and symbols with circles show detections of \frb. Multiple circles indicate multiple burst detections, except for Arecibo, which typically has multiple detections per observing session (a detailed analysis is left for a future paper). The black dashed lines show the VLA burst detections with simultaneous coverage at other telescopes. The solid black line shows the simultaneous burst detection at VLA and Arecibo.
\label{fig:multi}}
\end{center}
\end{figure*}

@@

% copied from the first paper
During the joint Arecibo-VLA campaign, Arecibo observed with the L-wide receiver, which has an observational frequency range of 1.15 to 1.73~GHz and a full width at half maximum beam size of 3.3 arcmin. The PUPPI pulsar backend was used to record total intensity spectra with time and frequency resolutions of 10.24~$\mu$s and 1.5625~MHz, respectively, and full Stokes polarization information. Each frequency channel was coherently dedispersed to 557 pc cm$^{-3}$, thereby eliminating intra-channel dispersion smearing. PUPPI covers a total of 800~MHz of bandwidth centred at 1380.78125~MHz, but only $\sim$ 620~MHz of this band is usable due to radio frequency interference and receiver sensitivity roll-off at the band edges.

@@

% copied from Laura's paper
Observations were conducted at an observing frequency of 4.6 to 5.1~GHz. The S60mm receiver has a system equivalent flux density of 18 Jy and a full-width half-max (FWHM) beam size of 2.4\arcmin at 4.85~GHz. Pulsar search mode data were recorded with the PFFTS backend. Total intensity spectra were recorded with a time resolution of 65.5~$\mu$s and a bandwidth of 500~MHz divided into 128 frequency channels. Note, the inter-channel DM smearing time for 560~pc cm$^{-3}$\ is $\sim$0.2 msec at 4.6 GHz.

@@

\subsubsection{Dynamic Spectra}
\label{sec:spec}
Figure \ref{fig:sgram} shows the spectrograms of all nine bursts detected by the VLA. We improved on the initial analysis presented in (CITE: LOC) by using a better calibration scheme and optimizing the image SNR over a fine DM grid. Table \ref{tab:spec} shows the improved burst parameters using this new scheme.

\begin{table*}
\caption{Properties of Bursts from \frb}
\centering
\begin{tabular}{lccccccc}
\hline
Date     & S$_{\rm{int}}$ & Image SNR & L$_{\rm{int}}$ & DM$_{\rm{opt}}$ & S$_{\rm{peak}}$ & Center & FWHM \\
(MJD)    & (mJy) &  & (ergs?) & (pc cm$^{-3})$ & (mJy?) & (GHz) & (MHz) \\ \hline
57623    & 258 & 38 & & 561 & 0.41             & 2.8 & 300 \\
57633.68 & 2000 & 179 & & 554 & 1.90           & 3.2 & 520 \\
57633.70 & 105 & 15 & & 559 & $>$0.188          & $<$2.5 & $>$350 \\
57638    & 65  & 12 & & 556 & 0.07             & 3.1 & 410 \\
57643    & 375 & 100 & & 560 & 0.39            & 2.8 & 520 \\
57645    & 38  & 13 & & 572 & 0.06             & 2.8 & 210 \\
57646    & 69  & 20 & & 555 & $>$0.16          & $<$2.5 & $>$400 \\
57648\tablenotemark{a}    & 97 & 25 & & 559 & 0.11 & 2.9 & 420 \\
57649    & 110 & 36 & & 552 & 0.07             & 2.9 & 880 \\ \hline
\end{tabular}
\tablenotetext{a}{Detected simultaneously with Arecibo between 1.15 and 1.73~GHz.}
\label{tab:spec}
\end{table*}

@@

The uncertainty in the peak DM measurement is defined by the DM sensitivity loss curve and the significance of the burst. We tested this model by simulating a burst with SNR$=30$ and measuring its detection significance as a function of DM. The simulation parameters were reproduced at the known DM with a precision of roughly 1 pc cm$^{-3}$. This suggests that the measured variation in DM for \frb is intrinsic to the bursts, rather than due to measurement uncertainty. This is consistent with the observation of intrinsic burst spectrotemporal structure that varies between bursts (CITE: WEIRD).

@@

\begin{figure*}[h!]
\begin{center}
 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.3\columnwidth]{spec_57623.png}
  \includegraphics[width=0.3\columnwidth]{spec_57633_scan7.png}
  \includegraphics[width=0.3\columnwidth]{spec_57633_scan13.png}
 \end{minipage}

@@

The burst spectra are generally characterized by a broad, Gaussian shape that with strong modulation between channels. We fit a Gaussian shape to each burst to estimate their characteristic width and peak flux, as summarized in Table \ref{tab:spec}. The channel-scale modulation is as high as 100\% (see \S \ref{sec:auto}) and is consistent with an exponential distribution. While this introduces a strong bias to the best-fit Gaussian, the typical burst spectral width has a characteristic width of 500~MHz.

@@

\subsubsection{Circular Polarization}
The VLA S-band recievers natively measure circular polarization, although observations did not include polarization calibration procedures. Crude constraints on circular polarization are possible by comparing the burst intensity in right and left-hand polarized data products. The apparent circular polarization fraction ($(RR-LL)/(RR+LL)$) for the most significant bursts are all less than 3\%. FRB121102 was located 2.3 arcmin away from pointing center, where systematic effects have been measured as large as 3\% (Perley et al 2016, VLA memo). We conclude that the S-band bursts seen by the VLA have a circular polarization of less than 3\%.

\subsection{Temporal Statistics}
Periodicity search...
Figure \ref{fig:ls} shows a Lomb-Scargle periodogram...
Lots of excess power corresponding to periods shorter than 1 hour...
Assuming periods with power peaks, we find clusters of bursts with same rotational phase...
Possible that we are seeing an underlying period with a very wide pulse phase window...

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{lombscargle}
\caption{Lomb-Scargle periodogram with 95\% limit on power from last observing campaign
\label{fig:ls}}
\end{center}
\end{figure}

Burst detections were made very inhomogenously through the roughly 63 hours of observing toward FRB 121102. In the first 30 hours of observing at S-band no bursts were detected, while nine bursts were detected in the last 27 hours of S-band observing. The data quality is high and RFI did not significantly impact sensitivity, so the inhomogeneous burst distribution is not an observational artifact, but rather an indication of a non-stationary burst detection probability. Assuming that the burst detection probability follows a Poisson distribution, the nondetection in the first half of S-band limits the FRB rate to $\rm{R}<0.1$ hour$^{-1}$ (95\% confidence limit). The mean detection rate for the last part of the campaign was $\rm{R}=0.3$ hour$^{-1}$.

Moreover, there is weak evidence that the \frb\ burst rate changes during the last part of the campaign. We modeled the event detection probability as a Poisson probability with a rate parameter that scales linearly with time (relative to first burst detection) as $\lambda = a + b * (\rm{MJD} - 57623)$. We then calculated the probability of measuring the observed event rate as $\prod_{i} P_i(R)$. Figure \ref{fig:rate} shows the ($a$, $b$) parameter space with contours of 50, 90, and 95\% confidence contours.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{event_rate_contours}
\caption{Rate versus rate change during last campaign.
\label{fig:rate}}
\end{center}
\end{figure}

\subsection{Brightness Distribution}
Table \ref{tab:spec} shows burst luminosities...
The significance of the nine bursts span more than an order of magnitude (12 to 179$\sigma$). This highlights the fact that the bursts have a relatively flat luminosity distribution.

Figure \ref{fig:logns} shows the flux distribution of the nine VLA bursts. Quantify flatness...

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{logns}
\caption{Log N -- Log SNR distribution of real-time detections of FRB 121102. **Update to luminosity distribution**
\label{fig:logns}}
\end{center}
\end{figure}

Compare luminosity distribution to flux distribution measured by \citep{2016ApJ...830...75V}...

@@

Discuss comparison FRB 121102 luminosity distribution to that of FRB population...
Imagine physical log N/log S with cut-offs and scattering can bias the intrinsic into the observed distribution (Macquart and Johnston)...

>>>>

This paper is part of a series that presents the first localization and unambiguous identification of an FRB host (CITE: LOC, OPT, EVN, RADIO CPART, XRAY, EFF, WEIRD). \frb, also known as the ``repeating FRB'', was first detected in November 2012 by the Arecibo Observatory \citep{2014ApJ...790..101S}. In mid 2015, new Arecibo observations revealed a series of bursts at the same DM and sky position demonstrating that FRBs are capable of repetition \citep{2016Natur.531..202S}. Beginning in August of 2015, we made the first of nine detections of \frb\ with the Very Large Array (CITE: LOC) and localized it with a precision of 0.1\arcsec. Deep radio and optical observing shows that \frb\ is unambiguously associated with a persistent radio and optical source at a redshift of 0.193 (CITE: OPT, EVN).

\frb\ has now been localized three orders of magnitude better than any other FRB and placed at a cosmological distance. Its lookback and luminosity distances are 746 and 972 Mpc \citep{planck15}, which are orders of magnitude larger than any other millisecond transient. This shows that FRBs are more luminous than any other millisecond radio transient, have a significant DM contribution from the IGM, and can be used to probe the IGM and their host galaxy. The promise implied by the first reported FRB \citep{2007Sci...318..777L} is now being realized.

The confirmation of a cosmological distance for \frb\ could have wide-ranging implications for the FRB population as a whole. However, it has not been demonstrated that \frb\ is representative of the overall FRB population. In fact, the repetition of its bursts is unique among all FRBs \citep{2015MNRAS.454..457P}, so it is natural to ask whether \frb\ is representative. An important first step is to demonstrate that the properties of \frb\ are consistent with the significant body of facts for the overall population \citep{2015MNRAS.451.3278M, 2016MPLA...3130013K}. The repeating nature of \frb\ provides us with several statistical tests we can use to test this connection.

@@

Given that FRBs are now known to be useful probes of the IGM, there is even more motivation to make new detections and localizations. The relatively faint counterpart to \frb\ argues that direct localization of the radio burst will continue to be the best way to find optical hosts to measure distances. Our multi-telescope constraints on burst spectra, measurement of host properties, burst rate estimates, and other properties will inform new strategies for finding FRBs.

@@

\begin{figure*}[t]
\begin{center}
\includegraphics[width=2\columnwidth]{timeline}
\caption{Summary of observing coverage and detections of \frb\ during the multi-telescope observing campaign in August and September 2016. Symbols show days with observations and circles highlight observations that detected bursts from \frb. Multiple circles indicate multiple burst detections, except for Arecibo, which typically has multiple detections per observing session (a detailed analysis is left for a future paper). The black dashed lines show the VLA burst detections with simultaneous coverage at other telescopes. The solid black line shows the simultaneous burst detection at VLA and Arecibo.
\label{fig:multi}}
\end{center}
\end{figure*}

@@

(copied from the first paper) 
During the joint Arecibo-VLA campaign, Arecibo observed with the L-wide receiver, which has an observational frequency range of 1.15 to 1.73~GHz and a full width at half maximum beam size of 3.3 arcmin. The PUPPI pulsar backend was used to record total intensity spectra with time and frequency resolutions of 10.24~$\mu$s and 1.5625~MHz, respectively, and full Stokes polarization information. Each frequency channel was coherently dedispersed to 557 pc cm$^{-3}$, thereby eliminating intra-channel dispersion smearing. PUPPI covers a total of 800~MHz of bandwidth centred at 1380.78125~MHz, but only $\sim$ 620~MHz of this band is usable due to radio frequency interference and receiver sensitivity roll-off at the band edges.

@@

(copied from Laura's paper)
Observations were conducted at an observing frequency of 4.6 to 5.1~GHz. The S60mm receiver has a system equivalent flux density of 18 Jy and a full-width half-max (FWHM) beam size of 2.4\arcmin at 4.85~GHz. Pulsar search mode data were recorded with the PFFTS backend. Total intensity spectra were recorded with a time resolution of 65.5~$\mu$s and a bandwidth of 500~MHz divided into 128 frequency channels. Note, the inter-channel DM smearing time for 560~pc cm$^{-3}$\ is $\sim$0.2 msec at 4.6 GHz.

@@

\subsubsection{Dynamic Spectra}
\label{sec:spec}
Figure \ref{fig:sgram} shows the spectrograms of all nine bursts detected by the VLA. We improved on the initial analysis presented in (CITE: LOC) by using a better calibration scheme and optimizing the detection signifnance by optimizing over a range of DM. Table \ref{tab:spec} shows the improved burst parameters using this new scheme.

\begin{table*}
\caption{Properties of Bursts from \frb}
\centering
\begin{tabular}{lccccccc}
\hline
Date     & S$_{\rm{int}}$ & Image SNR & L$_{\rm{int}}$ & DM$_{\rm{opt}}$ & S$_{\rm{peak}}$ & Center & FWHM \\
(MJD)    & (mJy) &  & ($10^{39}$\ erg) & (pc cm$^{-3})$ & (mJy?) & (GHz) & (MHz) \\ \hline
57623    & 258 & 38 & 1.5 & 561 & 0.41             & 2.8 & 300 \\
57633.68 & 2000 & 179 & 12 & 554 & 1.90           & 3.2 & 520 \\
57633.70\tablenotemark{a} & 105 & 15 & 0.6 & 559 & $>$0.188          & $<$2.5 & $>$350 \\
57638    & 65  & 12 & 0.4 & 556 & 0.07             & 3.1 & 410 \\
57643    & 375 & 100 & 2.1 & 560 & 0.39            & 2.8 & 520 \\
57645    & 38  & 13 & 0.2 & 572 & 0.06             & 2.8 & 210 \\
57646\tablenotemark{a}    & 69  & 20 & 0.4 & 555 & $>$0.16          & $<$2.5 & $>$400 \\
57648\tablenotemark{b}    & 97 & 25 & 0.6 & 559 & 0.11 & 2.9 & 420 \\
57649    & 110 & 36 & 0.6 & 552 & 0.07             & 2.9 & 880 \\ \hline
\end{tabular}
\tablenotetext{a}{Best-fit Gaussian is not centered in 3~GHz band, so spectral parameters are limits.}
\tablenotetext{b}{Detected simultaneously with Arecibo between 1.15 and 1.73~GHz.}
\label{tab:spec}
\end{table*}

@@

The uncertainty in the peak DM measurement is defined by the DM sensitivity loss curve and the significance of the burst. We tested the precision of the peak DM measuremet by simulating a burst with SNR$=30$ and measuring its detection significance as a function of DM. The simulation parameters were reproduced at the known DM with a precision of roughly 1 pc cm$^{-3}$. This suggests that the measured variation in DM for \frb is intrinsic to the bursts, rather than due to measurement uncertainty. This is consistent with the observation of intrinsic burst spectrotemporal structure that varies between bursts (CITE: WEIRD).

@@

\begin{figure*}[ht]
\begin{center}
 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.3\columnwidth]{spec_57623.png}
  \includegraphics[width=0.3\columnwidth]{spec_57633_scan7.png}
  \includegraphics[width=0.3\columnwidth]{spec_57633_scan13.png}
 \end{minipage}

@@

The burst spectra are generally characterized by a broad, Gaussian shape that with strong modulation between channels. We fit a Gaussian shape to each burst to estimate their characteristic width and peak flux, as summarized in Table \ref{tab:spec}. The channel-scale modulation is as high as 100\% (see \S \ref{sec:auto}) and is consistent with an exponential distribution. This introduces a strong bias to the best-fit Gaussian, but the models are reliable enough to show that the typical burst has a spectral width of 500~MHz.

@@

\subsubsection{Circular Polarization}
The VLA S-band recievers natively measure circular polarization, although for observing efficiency we chose not include polarization calibration procedures. Crude constraints on circular polarization are possible by comparing the burst intensity in right and left-hand polarized data products. The apparent circular polarization fraction ($(RR-LL)/(RR+LL)$) for the most significant bursts are all less than 3\%. FRB121102 was located 2.3 arcmin away from pointing center, where systematic effects have been measured as large as 3\% (Perley et al 2016, VLA memo). We conclude that the S-band bursts seen by the VLA have a circular polarization of less than 3\%.

\subsection{Temporal Statistics}
The VLA detected nine bursts during the month-long campaign in August--September 2016. The typical observation lasted about two hours and was sampled with a 5~ms cadence, so we used the Lomb-Scargle periodogram \citep{} to search for periodic structure over a wide range of timescales. The input "lightcurve" was mde as a one-dimensional array for all times with observations; times with a detection are set to 1 and other times are set to 0 \citep[see also]{2011MNRAS.417.1871P}. The lightcurve time resolution was set to {\color{red} 100~ms (check)} to keep the computation manageable. The 95\% confidence upper limit on periodic power is estimated by randomly selecting the 9 burst times from the same observing times. The limit is set as the 95\% highest point at each period (5th highest value in 100 simulations).

Figure \ref{fig:ls} shows the Lomb-Scargle periodogram for the bursts for periods from **xx to xx** s. No significant excess power is seen at long periods, but the periodogram shows a steady increase in power at shorter periods. The periodogram exceeds the 95\% power limit for periods from xx to xx s (**limited by binning time?**). This is consistent with the idea that the bursts imperfectly trace an underlying rotational period, as has been observed from many classes of neutron star. Examples include magnetars, which typically have wide pulse phase windows \citep{}, pulsars with glitches that change their periods \citep{}, and {\color{red} Maura? also refs for previous points would be welcome}. However, we can show with simulations that nine bursts drawn from a simple rotational would produce a strong, narrow peak in \ref{fig:ls}.

Assuming periods with power peaks, we find clusters of bursts with same rotational phase...

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{lombscargle}
\caption{Dots show the Lomb-Scargle periodogram of burst arrival times for \frb. The solid line show the 95\% confidence upper limit on the power expected from bursts with random arrival times. (**New version being compiled**)
\label{fig:ls}}
\end{center}
\end{figure}

Burst detections were made very inhomogenously through the larger (63 hours) observing campaign of FRB 121102. In the first 30 hours of observing at S-band no bursts were detected, while nine bursts were detected in the last 27 hours of S-band observing. The data quality is high and RFI did not significantly impact sensitivity, so the inhomogeneous burst distribution shows that the burst detection probability was not stationary. Assuming that the burst detection probability follows a Poisson distribution, the nondetection in the first half of S-band limits the FRB rate to $\rm{R}<0.1$ hour$^{-1}$ (95\% confidence limit). The mean detection rate for the last part of the campaign was $\rm{R}=0.3$ hour$^{-1}$.

Moreover, there is weak evidence that the \frb\ burst rate changes during the last part of the campaign. We modeled the event detection probability as a Poisson probability with a rate parameter that evolves linearly with time relative to first burst detection. We directly sampled the model by calculating the joint probability $\prod_{i} P_i$, where $P_i$ follows a Poisson probability with rate $\lambda = a + b * (\rm{MJD}_i - 57623)$. Figure \ref{fig:rate} shows that the probability distribution excludes a constant rate with $\sim$85\% confidence. This weak constraint is consistent with the broader trend seen by the VLA and Arecibo (CITE: WEIRD?).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{event_rate_contours}
\caption{Color scale and contours show the relative probability for a time-evolving Poisson detection probability for \frb. Contours show the 50, 90, and 95\% confidence contours on the rate $a$\ (bursts s$^{-1}$) and rate change $b$\ (bursts s$^{-1}$ day$^{-1}$) during the August--September 2016 campaign in which bursts were detected.
\label{fig:rate}}
\end{center}
\end{figure}

\subsection{Brightness Distribution}
Knowing the burst distance, we calculate their mean luminosity across a single 5~ms integration and the 1~GHz bandwidth (Table \ref{tab:spec}). The VLA has also shown for the first time that most bursts are contained within the 3~GHz band. That means that the mean S-band flux density can be converted to a luminosity in units of ergs with no assumptions about its spectral properties.

With a uniform sample of burst luminosities, we can use their distribution to infer more general properties for \frb. Figure \ref{fig:lumd} shows the luminosity distribution of the VLA bursts with a best-fit powerlaw model. Fitting a powerlaw model to the cumulative luminosity distribution gives an index of --0.5. The flatness of the distribution is also clear from range of significance of the nine bursts (12 to 179$\sigma$). All bursts are much brighter than the sensitivity limit (weakest is $\sim10\sigma$), so the distribution is not affected by a completeness limit.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{luminosity_disn}
\caption{Cumulative luminosity distribution and best-fit powerlaw model for the nine VLA bursts from \frb.
\label{fig:lumd}}
\end{center}
\end{figure}

@@

Discuss comparison FRB 121102 luminosity distribution to that of FRB population...

Imagine physical log N/log S with cut-offs and scattering can bias the intrinsic into the observed distribution (Macquart and Johnston)...

Simulate cosmic volume uniformly filled with \frb...
